`r params$project_title`

### 2020-12-11 reports/rmarkdown/03_integrated_datasets.Rmd
updates with manuscript template

metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 6 ] <- 6
zscore[ zscore < -6 ] <- -6
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)

#Compute sample-wise correlations
cors <- cor(zscore, method = "pearson")

#Graph construction  
k <- 5 # k for KNN
P <- 0.4# prunning threshold of SNN

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

@@

@@

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)

@@

```{r, echo = FALSE,fig.width=10,fig.height=4}
myletters <- letters

@@

  datasets <- factor(metadata$joint_clustering)
  # datasets <- datasets[ colSums(all_microbiome)>0 ]
  # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("CVL_v3",121) , rep("CVL_v2",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE}
groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
names(groupings) <- names(datasets_all_samples)[-1]

groupings_all_samples <- lapply(groupings,function(x){
  df <- data.frame(group=rep("NA",nrow(metadata)), row.names = rownames(metadata),stringsAsFactors = F)
  df[rownames(x),1] <- as.character(x[,1])
  return(df)
})
names(groupings_all_samples) <- names(groupings)

df <- as.data.frame(groupings_all_samples)
colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
metadata <- cbind(metadata,df)
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) )
datasets <- paste0(datasets, "_",c(metadata$louvain_tissue_V3,metadata$louvain_CVL_V3,metadata$louvain_CVL_V2))

datasets <- factor( datasets[ colSums(all_microbiome)>0 ])
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(2,2)
group_means <- sapply( levels(datasets), function(x) rowMeans( all_microbiome[,datasets==x] ) )
cors <- cor(group_means)
h <- hclust( as.dist( (1-cors)/2 ),method = "ward.D2" )
o <- h$order

par(mar=c(6,6,2,2))
image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
      main="correlation matrix",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(h$labels)),
        rev((0:(length(h$labels)-1) )/(length(h$labels)-1)),
            labels = h$labels[o], srt = 0, adj = c(1,.5), xpd = TRUE, cex=1)
text(  (0:(length(h$labels)-1) )/(length(h$labels)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(h$labels)),
            labels = h$labels[o], srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
add_letter("a")

par(mar=c(6,4,2,2))
plot(as.dendrogram(h),las=1,ylab="height")
add_letter("b")
```

**Figure 3.** Comparrisson among patient groups across datasets. **(a)** Correlation matrix across sample groups. **(b)** Hierachical clustering of sample groups.

```{r, echo = FALSE,fig.width=10,fig.height=12}
res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

@@

```{r, echo = FALSE,fig.width=8,fig.height=8}
metadata_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
                         "blood_cervix_v2","blood_cervix_v3",
                         "cervical_inflammation_v2","cervical_inflammation_v3",
                         "HIVstatus","Contraception",
                         "CT_v2","CT_v3","regular_partner",
                         "vaginal_douching_v2","vaginal_douching_v3",
                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3",
                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3","joint_clustering")

@@

# metadata$Contraception
#Antibiotics usage
for(i in metadata_parameters){
  empty_plot()
  text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
  for(j in c("louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")){
    x <- as.character(metadata[,i])
    x[is.na(x)] <- "NA"
    y <- as.character(metadata[,j])
    y[is.na(y)] <- "NA"
    x1 <- x [ (x!="NA") & (y!="NA") ]
    y1 <- y [ (x!="NA") & (y!="NA") ]
    plot_sankey( data.frame(x1,y1),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(13))
    text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
  }
}
```

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- bacteria_use

@@

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))

#Graph construction  
k <- 5 # k for KNN
P <- 0.2 # prunning threshold of SNN

@@

@@

# cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)

table(cl$membership)

@@

```{r, echo = FALSE,fig.width  = 10, fig.height=12}
mypar(3,6,mar=c(1,1,2,1))
for(j in sort(unique(cl$membership)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
  title(main=paste0("community",j))
}
```

@@

```{r, eval=T, echo = FALSE,fig.width=10, fig.height=10}
sample_order <- order(df$louvain_CVL_V3)

@@

@@

```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}
mypar(7,3,mar=c(3,5,2,2))

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

@@

#Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

@@

for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

>>>>

metadata[colnames(counts)[(colSums(counts == 0)>=3)],]

samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
metadata <- metadata[ samples_to_use , ]
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)

#Compute sample-wise correlations
cors <- cor(all_microbiome, method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

# cors <- cor(all_microbiome, method = "pearson")
# adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

#Graph construction  
k <- 10 # k for KNN
P <- 0.5# prunning threshold of SNN

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(adj,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

@@

jSNN[1:10,1:10]

k=20
P=0.1
snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }

for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }

snnm[ snnj <= ceiling(k*P) ] <- 0

@@

g <- graph_from_adjacency_matrix(snnm,mode = "undirected",diag = F,weighted = T)

@@

```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}

datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) ,
                      rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ,
                      rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))

groups <- factor( rep(metadata$joint_clustering,3) )
age <- rep(metadata$age,3)
work_time <- rep(metadata$sexwork_months,3)

datasets <- datasets[ colSums(all_microbiome)>0 ]
groups <- groups[ colSums(all_microbiome)>0 ]
age <- age[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

df <- data.frame(groups,datasets,age)

library(edgeR)
design <- model.matrix( ~ datasets * groups + age , data=df )
rownames(design) <- paste0(datasets,"_",colnames(all_microbiome))
colnames(all_microbiome) <- rownames(design)

y <- DGEList(counts = round(((2^all_microbiome)-1)*1000) , remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

#age
lrt <- glmLRT(fit,coef = 8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
plot(top$logFC,-log10(top$PValue))

#age
lrt <- glmLRT(fit,coef = 4:7)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

#age
lrt <- glmLRT(fit,coef = 9:12)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

#age
lrt <- glmLRT(fit,coef = 13:16)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]

```{r, echo = FALSE,fig.width=5,fig.height=5}
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0

mypar(3,1,mar=c(2,6,2,2))
barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Monbilicus",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardanerella",
         genes = c("BV","counts_16S"),
         clustering = metadata$joint_clustering,
         draw_mean_lines=F,col = pal)
```

```{r, eval=F,echo = FALSE,fig.width=10,fig.height=4}
myletters <- letters

@@

  datasets <- factor(metadata$joint_clustering)
  # datasets <- datasets[ colSums(all_microbiome)>0 ]
  # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
  NN <- min(table(datasets))

@@

```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, echo = FALSE}
groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
names(groupings) <- names(datasets_all_samples)[-1]
all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))

groupings_all_samples <- lapply(groupings,function(x){
  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
  df[rownames(x),1] <- as.character(x[,1])
  return(df)
})
names(groupings_all_samples) <- names(groupings)

df <- as.data.frame(groupings_all_samples)
colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
metadata <- cbind(metadata,df[rownames(metadata),])
```

@@

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))

datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

@@

```{r, eval=F,echo = FALSE,fig.width=10,fig.height=12}
res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

@@

```{r, echo = FALSE,fig.width=8,fig.height=8}
metadata_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
                         "blood_cervix_v2","blood_cervix_v3",
                         "cervical_inflammation_v2","cervical_inflammation_v3",
                         "HIVstatus","Contraception",
                         "CT_v2","CT_v3","regular_partner",
                         "vaginal_douching_v2","vaginal_douching_v3","joint_clustering")

@@

# metadata$Contraception
#Antibiotics usage
for(i in metadata_parameters){
  empty_plot()
  text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
  for(j in c("joint_clustering")){
    x <- as.character(metadata[,i])
    x[is.na(x)] <- "NA"
    y <- as.character(metadata[,j])
    y[is.na(y)] <- "NA"
    x1 <- x [ (x!="NA") & (y!="NA") ]
    y1 <- y [ (x!="NA") & (y!="NA") ]
    plot_sankey( data.frame(x1,y1),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(13))
    text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
  }
}
```

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T))
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- bacteria_use

@@

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=5,]))

#Graph construction  
k <- 5 # k for KNN
P <- 0.1 # prunning threshold of SNN

@@

# k=10
# P=0.1
snn <- dbscan::sNN(x = as.dist( (1-cors)/2 ), k = k, jp = FALSE, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
snnm[ snnj <= ceiling(k*P) ] <- 0

@@

# cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)
# 
table(cl$membership)

@@

```{r, echo = FALSE,fig.width  = 10, fig.height=12}
mypar(3,6,mar=c(1,1,2,1))
for(j in sort(unique(cl$membership)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
  title(main=paste0("community",j))
}
write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))

@@

```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}

@@

mypar(4,3,mar=c(3,5,2,2))

@@

```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}

mypar(4,3,mar=c(3,5,2,2))

# for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
#   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
#           "ASV_CVL_V3_normalized_batch_corrected",
#           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
#     temp_group <- as.character(df[,i])
#     temp_group[ is.na(temp_group) ] <- "NA"
#     temp_group[ temp_group == "" ] <- "NA"
#     temp_group <- factor(temp_group)
#     
#     barlist(data=module_means_per_dataset[[j]],
#             genes=rownames(module_means_per_dataset[[j]]),
#             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
#             clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
#   }
# }

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

@@

#Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = rev(sub("HALLMARK_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

par(mar=c(6,2,3,2))
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
# empty_plot(xaxs="i",yaxs="i",
#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

@@

for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = rev(sub("KEGG_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

@@

```{r, echo = FALSE,fig.width=10,fig.height=8}
write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
```
